\section{Introducing the Contiki Operating System}\label{sec:CONTIKI}

  \Contiki\ operating system can be seen as collection of processes.
 In the same way as the UNIX operating system is seen as collection of
 files and processes. However, files are not a necessary component, in
 fact \emph{Coffee File System (CFS)} is only needed for some specific
 applications which require a non-volatile storage abstraction. 
  The inter-process communication (IPC) in UNIX has not been implement
 in the very early releases, although Contiki processes wouldn't work
 without its event-driven IPC mechanism\footnote{\emph{Though it is very
 primitive comparing to the usual meaning of IPC as an acronym.}}, which also
 drives the core scheduler. There is no notion of users in \Contiki\,
 however, and at this point comparison with UNIX becomes apparently
 impossible. Nevertheless, such a high-contrast comparison will hopefully
 deliver a good picture of what \ContikiOS\ is and what it is not.

% check the date in the slides from WSN course

  The lead developer of \Contiki\, Adam Dunkels of Swedish Institute of
 Computer Science (SICS), has revolutionised embedded system world by
 implementing \emph{lwIP} in 2000 and later, in 2002, even further optimised
 \emph{uIP} stack. It is known that until \emph{lwIP}, no complete IP stack
 existed which could run on a microcontroller device \cite{dunkels03full}.

  The key technique utilised in implementing \Contiki\ and the \emph{uIP}
 stack is based on a very obscure behaviour which the C language's
 \texttt{switch/case} statement exhibits in certain context. Dunkels
 wrote a few papers \cite{dunkels05protothreads,dunkels06protothreads}
 describing how it had been achieved, and summarises it in his doctoral
 thesis \cite{dunkels07programming}. The threading technique has been
 named \emph{protothreads (PT)}. Processes in \Contiki\ are designed as
 an extension to the \emph{protothreads}. Another extension is called
 the \emph{protosockets}, as the name suggest it provides the abstraction
 of network sockets.

  Most of functionality of all three key elements is delivered by the use
 of C pre-processor macros. The \emph{protothreads} are platform
 independent\footnote{\emph{It is not completely true for Contiki, there
 are few additions which utilise facilities that some hardware platforms provide.}}
 and can be integrated into any C program by including only one header file.
 \emph{Protothreads} can be used with virtually any C compiler and,
 as described in \cite{dunkels06protothreads}, the overhead is rather
 insignificant, considering the powerful functionality; the only limitation
 is due to the use of \emph{\texttt{switch/case}} statements by the \emph{PT}
 macros, these control structures will lead to undefined behaviour of the
 code in the \emph{PT} context.

  \emph{Protosockets} are designed specifically for \emph{uIP} and therefore 
 cannot be included with just one header. The only limitation of the
 \emph{protosockets} is due to design decision, there is no UDP communication
 facility, only TCP is provided. The \emph{protosockets} greatly simplify the
 application code, since all necessary functions are provided, including
 handling of strings. Thought no packet flushing is currently possible, which
 may be desired in some application.

\subsection{Initial Development Phase}

  \Contiki\ source code includes a large set of stable drivers for various
 RCBs, which will be described in \ref{sec:hwintro}, including \emph{Zolertia Z1},
 \emph{Tmote Sky}, \emph{Atmel Raven} and \Chip{RF230}\emph{-based} boards,
 \emph{TI} \Chip{CC2430} and \emph{MSP430} as well as \emph{ARM Conrtex-MX}
 and many older devices, including \emph{Comondore 64} and \emph{Apple II}
 computers.

 Details regarding the \Chip{MC1322x} port \cite{links:contiki:port:mc1322x}
 are to follow in \ref{sec:MCX}, since \Chip{MC1322x} was chosen as the
 target platform for this project. However, the \RFA\ platform had been the
 first preference. Major work during the initial phase had been done to
 port the existing driver code for \Chip{RF230} transceiver to take the
 advantage of the single chip device. %This work had been described in the
 %interim report \cite{wmi:irep}.


  A considerably long period of work during the first semester had been
 towards porting the \ContikiOS\ to run on the \RFA\ chip. However, if
 the porting task was to be completed, the time it would have taken may
 have exceeded the allocated period of one academic year. Therefore it
 was understood that some reconsideration was required and two major
 solutions were proposed: either changing the development hardware or
 trying a different OS with the same hardware (\ref{sec:MCX}).

\subsubsection{Problems Encountered while Porting the Driver Code}

  It had been not a trivial task to compare a set of implementations
 of the driver code, nevertheless some common fragments of code had
 been identified. The interim report has included the details of what
 was achieved at the time \cite{wmi:irep}, though most of work in this
 area was taken as a preparation to the next phase.

  The log of commits made to the source code repository can be viewed at:

\footnotesize
\URL{https://github.com/contiki/contiki-rf2xx/commits/porting?author=errordeveloper}
\normalsize

  There had been two main implementations which were compared, although
 another two components had been looked at while working on this part
 of the project. \emph{Atmel} offers two software stack for \emph{6loWPAN}
 and two other packages for \emph{ZigBee} implementation \cite{atmel:avr2102,
 atmel:zbpro}. Only \emph{6loWPAN} was of interest to this project.

\begin{itemize}

	\item\emph{Atmel RUM} \cite{atmel:avr2070}
	\item\emph{Atmel MAC} \cite{atmel:avr2025}
	\item\emph{Contiki RF230 Driver}
	\item\emph{Contiki RF230BB Driver}

\end{itemize}

  After studying the source of these imlementations, it has become apparent
 that an earlier version of \emph{Atmel RUM} has been already ported to
 \emph{Contiki} and appears to be the base for the \Chip{AT86RF230} driver
 and the MCU core \Chip{ATmega1281}. The codename for that driver is
 \emph{RF230} and it's modified version is name \emph{RF230BB} (the
 suffix \emph{"BB"} stand for \emph{bare-bones}). \emph{RF230BB} is
 generally a stripped version of the first driver, it has been designed
 to suite up-to-date internals of \ContikiOS.

  The \emph{MAC} stack from \emph{Atmel} has been looked at and was
 considered  not suitable for it's license terms. The license can be
 found in the source code available form the \emph{Atmel} website
 \cite{atmel:avr2025}. Unlike the \emph{RUM} license, the \emph{MAC}
 license dictates a set of limitation on how the source code can be used, 
 this package cannot be distributed as part of an open-source project.
 Nevertherless, it appears that the \emph{Atmel MAC} implementation is
 generally more advanced (provides more functions) and has a clearer
 coding style and structure.
 
  A major issue arose, when the upstream \Contiki\ code tree has been
 modified by one of the developers\footnote{\emph{The commit history of
 the upstream AVR sub-tree can be viewed at:}
 \URL{https://github.com/contiki/contiki-mirror/commits/master/cpu/avr}},
 David Kopff, to provide an initial support for the \RFA\ chip. Those
 change were contradicting with what has been done while working on this
 project. Also it became apparent that no major improvements are going
 to be made to provide simpler facilities for porting \Contiki\ to new
 \emph{AVR} devices.
 Nevertheless, during a discussion on the mailing list, that has been
 agreed upon as generally a good direction for the development of the
 \emph{AVR} branch, thought had been seen as a remote target. Further
 study of the source code in the \emph{AVR} branch lead to a decision
 that a complete refactoring is necessary to provide an appropriate
 level of abstraction for the new \emph{AVR} device drivers to be
 designed. Also the separation of radio transceiver drivers, for use
 with other architecture branches, would be feasible. It is certainly
 a longer term task to leverage such changes and could not be achieved
 within the time allocated for this project.

  Another reason for reconsidering the goals of the project had been
 a possibility of encountering bugs in the driver code at some point
 closer to the project deadline, since the application code was yet to
 be written. In such situation, it would be difficult to find whether
 a bug is in the driver code or it is in the application code, hence
 the best idea is to design the application using a tested platform,
 where a minimum of low-level code would need to be written. And then,
 if desired, migrate it to other platforms, once it's all been approved
 as working correctly.

\subsubsection{Issue Tickets}

  The following bug tracker issue tickets had been open and are considered
 completed now, since this part of project has been suspended. This list
 is included here for refrence only.

 \begin{description} \em
	\IssueX{14} Compare Avaliable Driver Implementations
	\IssueX{15} AVR cross-compiler toolchain installation
	\IssueX{16} Porting Methodology
	\IssueX{20} Driver Code Licensing
	\IssueX{24} Early Compilation Errors
 \end{description}
