\newcommand{\Chip}[1]{\emph{\texttt{#1}}}
\newcommand{\Symb}[1]{\emph{\texttt{#1}}}
\newcommand{\MCX}[0]{\Chip{MC1322x}}
\newcommand{\RFA}[0]{\Chip{ATmega128RFA1}}
\newcommand{\Port}[1]{\emph{\texttt{Port #1}}}
\newcommand{\Comp}[1]{\emph{\texttt{#1}}}
\newcommand{\WPAN}[0]{\emph{LR-WPAN}}

\newcommand{\TrackerList}[0]{\subsubsection{\emph{Tracked Issues}}\begin{description}}
\newcommand{\TrackerEnd}[0]{\end{description}}

\newcommand{\Contiki}[0]{\emph{Contiki}}
\newcommand{\ContikiOS}[0]{\emph{Contiki OS}}
\newcommand{\TinyOS}[0]{\emph{TinyOS}}


\documentclass[a4paper]{report}

\usepackage{multirow}
\usepackage{graphicx}

%02:36 <kahrl> errordeveloper: http://physics.wm.edu/~norman/latexhints/conditional_macros.html
%01:37 <kahrl> so \makeatletter \ifx \PRINTING \@empty \renewcommand{\href}{FAKE_VERSION_OF_HREF} \fi \makeatother
%01:39 <kahrl> or \makeatletter \ifx \PRINTING \@empty \relax \else \renewcommand{\href}{FAKE_VERSION_OF_HREF} \fi \makeatother

% `latex '\def\NOHREF{}\input{foo.tex}'`

\ifdefined \NOHREF

	\usepackage{url}
	\newcommand{\URL}[1]{\[ \texttt{\emph{#1}} \]}
	%% \newcommand{\href}[2]{#2 (\texttt{\emph{\url{#1}}})} %% not needed probably
	
	\newcommand{\Href}[2]{{#2}} %% fake version for printing
\else

\usepackage{hyperref}
\newcommand{\Href}[2]{\href{#1}{#2}} %% use this for proper href
\newcommand{\URL}[1]{\[ \Href{#1}{\texttt{\emph{#1}}} \]}

\fi



%% use \Href  and for printing we use \cite{something}
%% so we end up with link to the URL and bibliography
%% reference number as well. If \Href is fake there
%% will be no URL in the text.
\newcommand{\AltRef}[3]{\Href{#2}{#1} \cite{#3}}



\newcommand{\Issue}[1]{\Href{http://wmi.new-synth.info/issues/#1}{Issue \##1}}
\newcommand{\IssueX}[1]{\item[\Href{http://wmi.new-synth.info/issues/#1}{\Issue{#1}}]:}

\newcommand{\TEP}[1]{\AltRef{\emph{TEP#1}}{http://www.tinyos.net/tinyos-2.x/doc/pdf/tep#1.pdf}{links:tinyos:tep#1}}

\newcommand{\RFC}[1]{\AltRef{\emph{RFC#1}}{http://tools.ietf.org/pdf/rfc#1.pdf}{ietf:rfc#1}}

%% usage:
%	\MAN{gcc}
%	\MAN{2, open}

%\newcommand{\MAN}[1]{\AltRef{\emph{\texttt{#1}}}{http://manpages.debian.net/cgi-bin/man.cgi?query=#1&sektion=0&format=pdf}{doc:linux:man:0:#1}}
%\newcommand{\MANX}[2]{\AltRef{\emph{\texttt{#2}}}{http://manpages.debian.net/cgi-bin/man.cgi?query=#2&sektion=#1&format=pdf}{doc:linux:man:#1:#2}}

\newcommand{\MAN}[1]{\AltRef{\emph{\texttt{#1}}}{http://manpages.debian.net/cgi-bin/man.cgi?query=#1&sektion=0&format=ascii}{doc:linux:man:0:#1}}
\newcommand{\MANX}[2]{\AltRef{\emph{\texttt{#2}}}{http://manpages.debian.net/cgi-bin/man.cgi?query=#2&sektion=#1&format=ascii}{doc:linux:man:#1:#2}}

%% For this commands we need to use \Href explicitly
%% so if Href is proper - the display the path and link it
%% otherwise - just display the path. Because there's no
%% reason to show various refences to this in the bibtex

\newcommand{\Blob}[1]{\Href{https://github.com/errordeveloper/contiki-wmi/blob/wmi-work/#1}{\texttt{#1}}}
\newcommand{\Tree}[1]{\Href{https://github.com/errordeveloper/contiki-wmi/tree/wmi-work/#1}{\texttt{#1}}}
\newcommand{\Blame}[1]{\Href{https://github.com/errordeveloper/contiki-wmi/blame/wmi-work/#1}{\texttt{#1}}}
\newcommand{\Logs}[1]{\Href{https://github.com/errordeveloper/contiki-wmi/commits/wmi-work/#1}{\texttt{#1}}}
\newcommand{\Contrib}[1]{\Href{https://github.com/errordeveloper/contiki-wmi/commits/wmi-work/#1?author=errordeveloper}{\texttt{#1}}}

\usepackage{tikz}
\usetikzlibrary{automata,positioning,shapes,arrows}

\tikzstyle{C} = [diamond,
			draw, fill=blue!20,
			text width=4.5em, text badly centered,
			node distance=3cm, inner sep=0pt]

\tikzstyle{F} = [rectangle,
			draw, fill=green!20,
			text width=5em, text centered,
			rounded corners, minimum height=2.5em]

\tikzstyle{E} = [ellipse,
			draw, fill=red!20,
			node distance=3cm, minimum height=3em]

\tikzstyle{L} = [draw, -latex']

\usepackage{pstricks-add}

%\DeclareGraphicsExtensions{.eps,.epsi,.pdf,.png,.jpg}

\title{WMI: Wireless Music Instruments\\ \emph{BSc Final Year Project} \\ Final Report}
\author{Ilya Dmitrichenko \\ \\ Department of Computing\\ London Metropolitan University}

\date{\today}



\begin{document}
\maketitle


\abstract
{

   This is the end of year report for a BSc degree project which had originally
   targeted a design of a complete wireless system for streaming of music control
   data collected from sensor devices and MIDI hardware to a host machine which
   would synthesise an audio signal to be delivered into loudspeakers or recorded
   on disk. Despite rather simple design idea, the research and development had
   taken extended time period and the emphasis of this report is on a variety of
   aspects related to wireless sensor network design and implementation. These
   aspects include: cross-platform operating system architecture, development
   hardware, application design and the abstraction layer requirements for
   interfacing hardware functions to the software as well as other miscellaneous
   subjects intersecting these areas. Two popular operating systems are presented
   in this report - \emph{Contiki} and \emph{TinyOS}. Two common microcontroller
   architectures had been used in the course of this project - 32-bit \emph{ARM}
   and 8-bit \emph{AVR}. \emph{IEEE 802.15.4} devices where considered as the
   only suitable radio harware option, due to the wide use and availability of
   the chips which implement this protocol. Another key assumption has been that
   the complice to the Internet Protocol (IP) is neccessary. A simple prototype
   has been implemented and a set of solutions proposed for a complete product
   design. Several advanced subjects for further work are also addressed in this
   report.

}

\tableofcontents
\listoffigures
%\listoftables

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\part{Introduction \& Research}
\chapter{General Background}

\include{bg/motivation}
\include{bg/organisation}
\include{bg/overview}

\part{Design \& Development} 
\chapter{Operating Systems for Wireless Sensor Networks}
\include{sw/node-rtos-intro}
\include{sw/node-contiki-intro}
\include{sw/node-tinyos-intro}
\chapter{Wireless Node Hardware}
\include{hw/node-intro}
\include{hw/node-rfa}
\include{hw/node-mcx}
\chapter{DSP Host System}
\include{hw/host-intro}
\include{sw/host-linux-intro}

\chapter{Application Development}
\include{sw/node-contiki-devel}
\include{sw/node-midi-intro}

\chapter{State Machines}
\include{sw/node-app-sm-adv-listener}
\include{sw/node-app-sm-adv-talker}

\chapter{System View}
\include{hw/host-integration}

\chapter{Looking Ahead}

  The major emphasis of this project had been in the area of development
 tools, platforms and code architecture for embedded software design.
 Though, originally, it was intended as an application development project,
 study of various alternatives of how the simple task can be achieved, lead
 to extended research in embedded operating systems and various programming
 tools, including: compiler toolchains, cross-platform integration techniques
 as well as the languages. Some of the most relevant aspects had been covered 
 in this report, leaving the unexplored depth behind.

  One of initial interest areas had been in the field of networking protocols
 for wireless sensing and control. Current best practice suggests that IP-enabled
 networking technology is most suitable for the global adoption, hence simple
 to integrate with various existing applications. It is rather difficult to
 imagine where, in the modern technology world, a non-IP network may be useful.
 Certainly, non-IP networking technology is available on the market, however if
 such devices are going to be used in any application, there are several limitations
 imposed on integration into various existing infrastructures. That implies
 physical connectivity as well as software support. One example is the popular
 \emph{ZigBee} family of protocols, which is gaining some popularity, thought
 no global scale of deployment has been observed. Not too long ago, there used
 to exist quite a few non-IP networking protocols, such as \emph{AppleTalk}
 and \emph{IPX}, for example. The majority of the published papers are oriented
 towards the use of IP-enabled networks for wireless sensing, just a few praise
 proprietary technologies.

  The project, which had been described here, has achieved its most
 fundamental goals, i.e. it has been proven that the MIDI data is
 handled as expected on a \emph{P802.15.4/6loWPAN} network and the
 platform had been established for future experimentation.
 The bigger project is now to commence.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{acm}

\bibliography{../main,../../contiki/dunkels,../../timesync/papers,../../tinyos/tep}

\end{document}
