\emph{Over View of Contiki Operating System} \label{sec:contiki:general}

  \emph{Contiki} operating system can be seen as collection of processes.
 In the same way as the UNIX operating system is seen as collection of
 files and processes. However, files are not a necessary component, in
 fact \emph{Coffee File System (CFS)} is only needed for some specific
 applications which require a non-volatile storage abstraction. 
  Inter-process communication in UNIX has not been implement in the very
 early releases, though Contiki processes wouldn't work without its  
 event-driven IPC mechanism\footnote{\emph{Though it is very primitive
 comparing to the usual meaning of IPC acronym.}}, which also drives the
 core scheduler. There is no notion of users in \emph{Contiki}, however,
 and at this point comparison with UNIX becomes apparently impossible.
 Nevertheless, such high-contrast comparison will hopefully deliver
 a good picture of what \emph{Contiki} is and what it is not.

% check the date in the slides from WSN course

  The lead developer of \emph{Contiki}, Adam Dunkels of Swedish Institute
 of Computer Science (SICS), has revolutionised embedded system world by
 implementing \emph{lwIP} in 200X and later, in 200Y, even further optimised
 \emph{uIP} stack. It is known that until \emph{lwIP}, no complete IP stack
 existed which could run on a microcontroller device \cite{dunkels03full}.

  The key technique utilised in \emph{Contiki} and the \emph{uIP} stack
 is based on a very obscure behaviour which the C language's \emph{
 texttt{switch/case}} statement possesses. Dunkels wrote a few papers
 \cite{dunkels05protothreads,dunkels06protothreads} describing how it
 had been implemented, and summarises it in his doctoral thesis
 \cite{dunkels07programming}. The threading technique has been named
 \emph{protothreads (PT)}. Processes in \emph{Contiki} are designed as
 an extension to the \emph{protothreads}. Another extension is called
 the \emph{protosockets}.

  Most of functionality of all three key elements is delivered by the use
 of C pre-processor macros. The \emph{protothreads} are platform
 independent\footnote{\emph{It is not true for Contiki, there are few
 additions which utilise facilities that some platform provide.}} and can
 be integrated into any C program by including only one header file.
 \emph{Protothreads} can be used with virtually any C compiler toolchain and
 as described in \cite{dunkels06protothreads}, the overhead is rather
 insignificant, considering the powerful functionality; the only limitation
 is due to the use of \emph{\texttt{switch/case}} statements by the \emph{PT}
 macros, these control structures will lead to undefined behaviour of the
 code in the \emph{PT} context.

  \emph{Protosockets} are designed specifically for \emph{uIP} and therefore 
 cannot be included with just one header. The only limitation of the
 \emph{protosockets} is due to design decision, there is no UDP communication
 facility, only TCP is provided. The \emph{protosockets} greatly simplify the
 application code, since all necessary functions are provided, including
 handling of strings. Thought no packet flushing is currently possible, which
 may be desired in some application.
 


   
