\section{DSP Host Hardware}

  A set of hardware platforms suitable for DSP host were considered.
 Among those are the following:
 	\begin{itemize}
		\item \emph{Analog Devices SHARC}
		\item \emph{ARM:} \begin{itemize}
		\item \emph{Marvell Sheeva}
		\item \emph{Texas Instruments OMAP}
		\item \emph{Freescale i.MX}
		\end{itemize}
	\end{itemize}
 All of these processor architectures are fully supported and widely
 used in embedded systems. Various other architecture families were
 considered, but found rather inappropriate due to the price range
 dictated by the target of this project. Multi-core \emph{MISP64}
 devices by \emph{NetLogic} \cite{netlogic:mips64:multicore}, for
 example, have great computational capacity, thought these are too
 expensive for this particular application.

\subsection{x86-based Embedded Devices}

  The list above does not include the \emph{x86}-based CPUs due to a
 fact that it had been very difficult to find a target board using
 Intel or any \emph{x86} CPU from other vendors. There are many boards
 from a large variety of manufacturers, hence the selection process
 becomes particularly time consuming. Quality evaluation would also
 be necessary\footnote{\emph{Due to the scale of production in this
 market, there is a great chance to obtain a device with a failure.}},
 when looking at \emph{x86} devices. Although, there are boards that
 would match the low-power and small form-factor requirements, most
 of these are with a handful of peripherals, for example low quality
 audio and video interfaces and, if these are not present, the board
 may have two or more RS-232 connectors. Another aspect of \emph{x86}
 system design is that it comes with legacy \emph{BIOS} technology,
 while most of \emph{ARM} systems, for example, have rather more
 flexible facilities for booloader and set-up. Also in the class of
 \emph{single board computers (SBC)} there are many boards which do
 not inlude physical connectors for above mentioned peripherals,
 however isuch boards are designed for industrial use require a
 specialised chasis \cite{links:linuxfordevices:guide}.

  Nevertheless, an \emph{x86} machine has been used for most of the
 development work on this project, that is for a rather transparent
 reason.

%% add to bibtex file:
% http://www.linuxfordevices.com/c/a/Linux-For-Devices-Articles/Single-Board-Computer-SBC-Quick-Reference-Guide/

\subsection{OMAP}

  The \emph{OMAP} application processors form \emph{TI} has become
 popular in the open-source community, and in fact \emph{TI} promotes
 Linux and Android as most suitable operating systems for this platform.
 The \emph{OMAP} architecture is based around an \emph{ARM} processor
 (there are various models with different versions of \emph{ARM} core)
 and \emph{TI C64x} DSP block. However, there is a little problem
 associated with how the DSP unit is integrated within the SoC.
 Very limited documentation is provided on how it can be used in the
 \emph{Linux} environment \cite{ti:omap:wiki:dsp}.

 One most outstanding development platform that uses an \emph{OMAP}
 chips with a dual-core \emph{Cortex-A8} CPU clocked at 1GHz - is the
 \emph{PandaBoard} \cite{ti:omap:wiki:pb}. However, it is available
 for back-order only, the maker is producing these boards on demand
 and the lead time would be at least one month. Otherwise this board
 would have been purchased despite the fact that DSP unit would be
 difficult to utilised.

 %http://www.omappedia.org/wiki/DSPBridge_Project
 %http://www.omappedia.org/wiki/PandaBoard

\subsection{Other ARM CPUs}

  \emph{ARM} CPUs are produced by almost any major semiconductor firm,
 so do \emph{Freescale} and \emph{Marvell}. Not all of those companies
 make very high performance \emph{ARM} chips, i.e. clocked near to 1GHz,
 and few make multi-core CPUs. And only some CPUs have an FPU.
 \emph{ARM} floating point instruction set is know as \emph{VFP}
 ("Vector Floating Point"). There are different versions of \emph{VFP},
 though in this context this wasn't a concern.

\subsubsection{Marvell Sheeva}

  \emph{Marvell} has originally started offering \emph{ARM} processors
 since their purchase of Intel's \emph{PXA} devision. Now \emph{Marvell}
 produces a series of high-performance \emph{ARM} chips, most of which
 are multi-core. \emph{Sheeva} is the brand name for these SoC devices.
 The clock frequency ranges from 800 MHz to 1.6 GHz, gigabit Ethernet
 is one of the outstanding features, since \emph{Marvell} specialises
 in the networking and storage IC market. As mentioned above, most
 important feature of \emph{ARM} chips that is necessary for the design
 of DSP host for this project is \emph{VFP}. Some of \emph{ARMADA}
 SoCs include \emph{VFP}, namely \emph{ARMADA XP} series,
 \emph{ARMADA 510} and \emph{610} \cite{links:marvell:armada}.
 \emph{Kirkwood} series are not featuring \emph{VFP}, thought there
 are some outstanding embedded development platforms available.

  \emph{Plug Commputer} (also know as \emph{Seeva Plug}) is small, yet
 very powerful computer in a form-factor of wall socket DC adaptor
 \cite{links:marvell:plug,links:plugcomp:homepage}. There many exciting
 application where these devices could be the best fit, thought due to
 above mentioned lack of \emph{VFP}, this CPU is not well suited as the
 DSP host.

% Marvell Semiconductor, Inc. - ARMAD Processor Family
% http://www.marvell.com/products/processors/armada.html
% http://www.marvell.com/platforms/plug_computer/
% http://www.plugcomputer.org/

\subsubsection{Freescale i.MX}

 \emph{Freescale Semiconducters} offers a range of \emph{ARM} processors
 branded emph{i.MX} \cite{links:freescale:imx}. The datasheets had been
 examined to find out which chips include floating point unit. Among the
 mid-range, \Chip{i.MX31} devices (532 MHz ARM1136JF-S) appear to have
 support for floating point instructions (\emph{VFP}), however the clock
 rate of this SoC would suite only a limited number of applications
 \cite{links:freescale:imx31}. In the upper-range of \emph{Freescale i.MX}
 solutions, there single-core \emph{Conrtex-A8} devices with clock rates
 up to a GHz \cite{links:freescale:imx5}, namely the \Chip{i.MX535}  and
 \Chip{i.MX538}. The second SoC has better video coding and storage
 capabilities.

 I has been noted that a simple development board for \Chip{i.MX535} can
 be purchased for a relatively low price\foonote{\emph{There two basic
 options: one for 150 US dollars, and another for 200 with a touch-screen.}}.

@MISC{links:freescale:imx,
	TITLE = "\href{http://cache.freescale.com/files/32bit/doc/brochure/FLYRIMXPRDCMPR.pdf}
		{{i.MX Product Family Overview}}",
	AUTHOR = {{Freescale Semiconductors}}
}


@MANUAL{links:freescale:imx31,
	TITLE = "\href{http://www.freescale.com/files/32bit/doc/ref_manual/MCIMX31RM.pdf}
		{{i.MX32 Reference Manual}}",
	AUTHOR = {{Freescale Semiconductors}}
}

@MANUAL{links:freescale:imx5,
	TITLE = "\href{http://cache.freescale.com/files/32bit/doc/fact_sheet/IMX5CNFS.pdf}
		{{i.MX5 Fact Sheet}}",
	AUTHOR = {{Freescale Semiconductors}}
}


\subsection{SHARC}

  \emph{Analog Device} is a well-known manufacturer of DSP chips, this
 architercture is different from \emph{TI C64x} and the signal processing
 instructions are handled directly on core. The clock frequencies are
 not being the main performance factor for the \emph{SHARC} device and
 range between 300 and 600 MHz. Linux has been a popular platform for
 advanced \emph{SHARC}-based systems appearing in various application
 markets \cite{links:adi:sharc}. It should noted that the advertisment
 of \emph{SHARC} chips provides less details then products from other
 vendors, hence an extensive evalution would be required for designing
 a system with \emph{SHARC}.

% http://www.analog.com/static/imported-files/product_highlights/SHARC_Proc_Family_(B)_Final.pdf

  \emph{SRARC} processors are commonly used in professional audio
 equipment. One very outstanding product has been released by
 a new British company \emph{Dark Matter Audio}, the device called
 \emph{DMA1} \cite{links:dma1} has been released this year.
 It is an advanced system for interactive music performance with
 very novel design features and powerful \emph{SHARC} DSP engine.
 According to the product announcement it runs Linux OS and also
 uses an \emph{ARM} chip for graphical user interface. It would be
 possible to integrate \emph{DMA1} with current prototype system
 developed in the course of this project. The vendor has provided
 an SDK, thought a request has been sent to obtain more specific
 details needed to design this add-on solution appropriately.


  \emph{SHARC} development hardware is mostly available from the
 silicon vendor itself and very few third-party boards could be found.
 Since this architecture is more specialised, the prices of the 
 development boards are not as low as some of the \emph{ARM} boards,
 although the vendor offers a comprehensive free library of board
 design file \cite{links:adi:freepcb}.

% http://blackfin.uclinux.org/gf/project/stamp/frs

